### 2014-02-05 trunk/ feature-selection-high-recall/FeatureSelectionProject/doc/fs_hr/background.tex
Intermediate update
Update

In this section, we briefly define the task of binary classification
along with standard definitions of performance metrics we may wish 
to optimize.  We then follow this by a discussion of feature selection and
existing criteria proposed in the literature.

In the binary classification task, we assume we are given data $D = \{
(\vec{x}^d, y^d) \}$ consisting of pairs of real-valued raw 
input vectors $\vec{x}^d \in \mathbb{R}^n$
of length $n$ (e.g., the results of $n$ different medical tests) and
\emph{actual} binary class label $y^d \in \{ 0 (\mathrm{false}), 1 (\mathrm{true}) \}$ (where
we often write F for false and T for true).
A binary classifier is a function $C: \vec{x}^d \to y^d$ such that given
a new unlabeled raw feature vector, $C(\vec{x}^d)$ produces a \emph{predicted}
classification.

Given a trained classifier $C$ and a dataset $D$, we can build the 
well-known contingency table
\begin{center}
\begin{tabular}{l|l|l|} 
\multicolumn{1}{l}{} &  \multicolumn{1}{l}{Actual T} & \multicolumn{1}{l}{Actual F} \\ \cline{2-3}
Predicted T & TP & FP \\ \cline{2-3}
Predicted F & FN & TN \\ \cline{2-3}
%Predicted T & TP & FP & \#$T_P$ \\ \cline{2-3}
%Predicted F & FN & TN & \#$F_P$ \\ \cline{2-3}
%          &   & \#$T_A$ & \#$F_A$ & \\ \cline{2-3}
\end{tabular}
\end{center}
where the four entries represent the counts of true positives (TP), false positives (FP), false negatives (FN)
and true negatives (TN) and sum to the 
total amount of data (i.e., $\mathrm{TP} + \mathrm{FP} + \mathrm{FN} + \mathrm{TN} = |D|$).
Each table entry represents the count of data for which the respective row matched the 
predicted classification $C(\vec{x}^d)$ and the respective column matched the actual label $y^d$.
Given these definitions, we can easily define 
\begin{align*}
\textrm{Accuracy}  & = \frac{TP + TN}{TP + FP + FN + TN}
%\textrm{Precision} & = \frac{TP}{TP + FP}
\textrm{Recall}    & = \frac{TP}{TP + FN}
\end{align*}
where accuracy represents the overall fraction of correct classification, but recall
represents the overall fraction of true labeled data that has been classified as
true (i.e., recalled).  For somewhat rare events such as medical diagnosis of cancer,
we certainly care about accuracy, but we may also want to place additional emphasis
on recall performance so as to avoid the occurrence of false negatives (cases of cancer
that were missed by the classifier).  Of course false positives are also a problem,
but additional tests would rule these out and hence not as critical of a classification
failure as missing a potential cancer diagnosis.

Previously we did not specify exactly how $C(\vec{x}^d)$ learns from the raw data
$\vec{x}^d$.  In general, practitioners often process the raw input vectors
$\vec{x}^d$ into a subset of features that we'll denote $f_1,\ldots,f_k$.  Whereas
the raw input may consist of the results of individual tests, a feature $f_i \in \mathbb{R}$
may represent some nonlinear function of one or more tests deemed to be relevant to
the classification task.  Hence we might more appropriately write a classifier
as $C(\vec{x}^d,f_1,\ldots,f_k)$ to represent the raw input data and the features
of the data that the classifier may use.
In the case of high-dimensional data or few data samples,
we may wish to limit the set of features generated to improve classifier performance
and this is the task of feature selection --- select $\{f_1,\ldots,f_k\}$ to optimize
performance.  There are already a variety of feature selection methods that we outline
next.

>>>>

In this section, we briefly define the task of binary classification
along with standard definitions of performance metrics we may wish 
to optimize.  We then follow this by a discussion of feature selection.
% and existing criteria proposed in the literature.

In the binary classification task, we assume we are given data $D = \{
(\vec{x}^d, y^d) \}$ consisting of pairs of real-valued raw 
input vectors $\vec{x}^d \in \mathbb{R}^n$
of length $n$ (e.g., the results of $n$ different medical tests) and
\emph{actual} binary class label $y^d \in \{ 0 (\mathrm{ false}), 1 (\mathrm{ true}) \}$ (where
we often write F for false and T for true).
A binary classifier is a function $C: \vec{x}^d \to y^d$ such that given
a new unlabeled raw feature vector, $C(\vec{x}^d)$ produces a \emph{predicted}
classification.

Given a trained classifier $C$ and a dataset $D$, we can build the 
well-known contingency table
\begin{center}
\begin{tabular}{l|l|l|} 
\multicolumn{1}{l}{} &  \multicolumn{1}{l}{Actual T} & \multicolumn{1}{l}{Actual F} \\ \cline{2-3}
Predicted T & TP & FP \\ \cline{2-3}
Predicted F & FN & TN \\ \cline{2-3}
%Predicted T & TP & FP & \#$T_P$ \\ \cline{2-3}
%Predicted F & FN & TN & \#$F_P$ \\ \cline{2-3}
%          &   & \#$T_A$ & \#$F_A$ & \\ \cline{2-3}
\end{tabular}
\end{center}
where the four entries represent the counts of true positives (TP), false positives (FP), false negatives (FN)
and true negatives (TN) and sum to the 
total amount of data (i.e., $\mathrm{TP} + \mathrm{FP} + \mathrm{FN} + \mathrm{TN} = |D|$).
Each table entry represents the count of data for which the respective row matched the 
predicted classification $C(\vec{x}^d)$ and the respective column matched the actual label $y^d$.
Given these definitions, we can easily define 
\begin{align*}
\textrm{Accuracy}  & = \frac{TP + TN}{TP + FP + FN + TN} \\
%\textrm{Precision} & = \frac{TP}{TP + FP}
\textrm{Recall}    & = \frac{TP}{TP + FN}
\end{align*}
where accuracy represents the overall fraction of correct classification, but recall
represents the overall fraction of true labeled data that has been classified as
true (i.e., recalled).  For somewhat rare events such as medical diagnosis of cancer,
we certainly care about accuracy, but we may also want to place additional emphasis
on recall performance so as to avoid the occurrence of false negatives (cases of cancer
that were missed by the classifier).  Of course false positives are also a problem,
but additional tests would rule these out and hence not as critical of a classification
failure as missing a potential cancer diagnosis.

Previously we did not specify exactly how $C(\vec{x}^d)$ learns from the raw data
$\vec{x}^d$.  In general, practitioners often process the raw input vectors
$\vec{x}^d$ into a subset of features that we'll denote $F_k = \{ f_1,\ldots,f_k \}$.  Whereas
the raw input may consist of the results of individual tests, a feature $f_i \in \mathbb{R}$
may represent some nonlinear function of one or more tests deemed to be relevant to
the classification task.  Hence we might more appropriately write a classifier
as $C(\vec{x}^d,F_k)$ to represent the raw input data and the features
of the data that the classifier may use.
In the case of high-dimensional data or few data samples,
we may wish to limit the set of features generated to improve classifier performance
and this is the task of feature selection --- select $F_k$ to optimize
performance.  There are already a variety of feature selection methods that we outline
next.

### 2014-02-05 trunk/ feature-selection-high-recall/FeatureSelectionProject/doc/fs_hr/empirical.tex
Update

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table*}[tbp!]
%\vspace{-1mm}
\centering
{\small 
\begin{tabular}{|l|r|r|r|r|l|} \hline
Dataset & \# Features & \# Data & \# Features/\# Data & \% True Labels & Feature Type\\ \hline \hline 
Breast Cancer(BC)	& 9	& 699	& 0.013	& 34 \%	& All Numeric\\ \hline
Diabetes(D)        & 8	& 768	& 0.010	& 35 \%	& All Numeric\\ \hline
Heart Statlog(HS)	& 13	& 270	& 0.048	& 44 \%	& All Numeric\\ \hline
Spect(S)	        & 22	& 80	& 0.275 & 33 \%	& Categorical\\ \hline
Vote(V)	        & 16	& 435	& 0.037	& 39 \%	& Categorical\\ \hline
Newsgroup(NG)	& 500	& 1963	& 0.255	& 49 \%	& All Binary\\ \hline
Horse Colic(HC)	& 22	& 368	& 0.060	& 37 \%	& Categorical (16) and numeric (6)\\ \hline
Credit-American(CR)	& 15	& 690	& 0.022	& 44 \%	& Categorical(9) and numeric (6)\\ \hline
Credit-German(GC)	& 20	& 1000	& 0.020 & 30 \%	& Categorical (13) and Numeric (7)\\ \hline
Hepatitis(H)	& 19	& 155	& 0.123	& 21 \%	& Categorical (13) and Numeric (6)\\ \hline
Ionosphere(I)	& 34	& 351	& 0.097	& 36 \%	& All Numeric\\ \hline
%KR-vs-KP	& 36	& 3196	& 0.011	& 48 \%	& All Categorical\\ \hline
%Labor	        & 16	& 57	& 0.281	& 35 \%	& Categorical (8) and Numeric (8)\\ \hline
%Mushroom	& 22	& 8124	& 0.003	& 48 \%	& All Categorical\\ \hline
%Sick	        & 29	& 3772	& 0.008	&  6 \%	& Categorical (23) and Numeric (6)\\ \hline
%Sonar	        & 60	& 208	& 0.288 & 47 \%	& All Numeric\\ \hline
\end{tabular}}
\vspace{-1mm}
\caption{\footnotesize Descriptive statistics of the various datasets evaluated in this work.}
\label{table:stats}
\vspace{-4mm}
\end{table*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[tbp!]
\centering
\includegraphics[width=0.375\textwidth]{./figures/linegraphs/SpectLogisticRegressionRecall.pdf}
\includegraphics[width=0.375\textwidth]{./figures/linegraphs/SpectNaiveBayesRecall.pdf}\\
\includegraphics[width=0.375\textwidth]{./figures/linegraphs/NewsgroupLogisticRegressionRecall.pdf}
\includegraphics[width=0.375\textwidth]{./figures/linegraphs/NewsgroupNaiveBayesRecall.pdf}
%\includegraphics[width=0.375\textwidth]{./figures/linegraphs/SpectLogisticRegressionRecall2.pdf}
%\includegraphics[width=0.375\textwidth]{./figures/linegraphs/SpectNaiveBayesRecall2.pdf}\\
%\includegraphics[width=0.375\textwidth]{./figures/linegraphs/NewsgroupLogisticRegressionRecall2.pdf}
%\includegraphics[width=0.375\textwidth]{./figures/linegraphs/NewsgroupNaiveBayesRecall2.pdf}
\caption{\footnotesize Performance distribution of various feature selection algorithms for two datasets and classifier types at each stage of feature selection.  The conjunctive voting methods do best overall in these datasets.}
\label{fig:perf_vs_fs_alg}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

In this section, we conduct experiments to study the feature selection
performances of several well known methods such as Symmetrical
Uncertainty Rank (SUR), Gain Ratio Rank (GRR), Mutual Information Rank
(MIR), Correlation Based Rank (CR), Conditional Entropy Rank (CER)~\cite{guyon_jmlr03}, Correlation-based Feature Subset Selection (CS)~\cite{Hall1998}, Reliff (RF)~\cite{Robnik-Sikonja2003} , Minimum Redundancy Maximum Relevancem(MRMR)~\cite{peng2005} and our proposed methods:  Disjunction Voting Expectation(DVE) and  Disjunction Voting likelihood (DVL), Conjunction Voting Expectation (CVE), Conjunction Voting likelihood (CVL). The first five methods are variable rank methods and the last seven are subset rank methods. The first five methods are variable rank methods that select variables by ranking them with some metric and the last seven are subset rank methods that assess subsets of variables and consider previous selections to decide the next selection.

@@

Next we examine Figure~\ref{fig:perf_vs_fs_alg}, where we evaluate the
performance distribution of various feature selection algorithms with
samples taken for each dataset, classifier type, and stage of feature
selection. This graph shows that our CVL high recall feature selection approach 
improves recall with the most improvement noticeable when there is a high
feature to data ratio such as in the Newsgroup (NG) and Spect (S) data
sets.  
%We conjecture the reason for this stems from the strong agreement of
%features provided by the conjunctive voting scheme that helps reduce noise
%during feature selection process.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table*}[tbp!]
%\vspace{-1mm}
\centering
{\small 
\begin{tabular}{|l|r|r|r|r|l|} \hline
Dataset & \# Features & \# Data & \# Features/\# Data & \% True Labels & Feature Type\\ \hline \hline 
Breast Cancer (BC)	& 9	& 699	& 0.013	& 34 \%	& All Numeric\\ \hline
Diabetes (D)        & 8	& 768	& 0.010	& 35 \%	& All Numeric\\ \hline
Heart Statlog (HS)	& 13	& 270	& 0.048	& 44 \%	& All Numeric\\ \hline
Spect (S)	        & 22	& 80	& 0.275 & 33 \%	& Categorical\\ \hline
Vote (V)	        & 16	& 435	& 0.037	& 39 \%	& Categorical\\ \hline
Newsgroup (NG)	& 500	& 1963	& 0.255	& 49 \%	& All Binary\\ \hline
Horse Colic (HC)	& 22	& 368	& 0.060	& 37 \%	& Categorical (16) and numeric (6)\\ \hline
Credit-American (CR)	& 15	& 690	& 0.022	& 44 \%	& Categorical(9) and numeric (6)\\ \hline
Credit-German (GC)	& 20	& 1000	& 0.020 & 30 \%	& Categorical (13) and Numeric (7)\\ \hline
Hepatitis (H)	& 19	& 155	& 0.123	& 21 \%	& Categorical (13) and Numeric (6)\\ \hline
Ionosphere (I)	& 34	& 351	& 0.097	& 36 \%	& All Numeric\\ \hline
%KR-vs-KP	& 36	& 3196	& 0.011	& 48 \%	& All Categorical\\ \hline
%Labor	        & 16	& 57	& 0.281	& 35 \%	& Categorical (8) and Numeric (8)\\ \hline
%Mushroom	& 22	& 8124	& 0.003	& 48 \%	& All Categorical\\ \hline
%Sick	        & 29	& 3772	& 0.008	&  6 \%	& Categorical (23) and Numeric (6)\\ \hline
%Sonar	        & 60	& 208	& 0.288 & 47 \%	& All Numeric\\ \hline
\end{tabular}}
\vspace{-2mm}
\caption{\footnotesize Descriptive statistics of the various datasets evaluated in this work.}
\label{table:stats}
\vspace{-3mm}
\end{table*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[tbp!]
\centering
\includegraphics[width=0.375\textwidth]{./figures/linegraphs/SpectLogisticRegressionRecall.pdf}
\includegraphics[width=0.375\textwidth]{./figures/linegraphs/SpectNaiveBayesRecall.pdf}\\
\includegraphics[width=0.375\textwidth]{./figures/linegraphs/NewsgroupLogisticRegressionRecall.pdf}
\includegraphics[width=0.375\textwidth]{./figures/linegraphs/NewsgroupNaiveBayesRecall.pdf}
%\includegraphics[width=0.375\textwidth]{./figures/linegraphs/SpectLogisticRegressionRecall2.pdf}
%\includegraphics[width=0.375\textwidth]{./figures/linegraphs/SpectNaiveBayesRecall2.pdf}\\
%\includegraphics[width=0.375\textwidth]{./figures/linegraphs/NewsgroupLogisticRegressionRecall2.pdf}
%\includegraphics[width=0.375\textwidth]{./figures/linegraphs/NewsgroupNaiveBayesRecall2.pdf}
\caption{\footnotesize Performance distribution of various feature selection algorithms for two datasets and classifier types at each stage of feature selection.  The conjunctive voting methods (CVE and CVL)
perform well (and reliably) across all four combinations of classifier and dataset.}
\label{fig:perf_vs_fs_alg}
\vspace{-3mm}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

In this section, we conduct experiments to study the feature selection
performances of several well known methods such as Symmetrical
Uncertainty Rank (SUR), Gain Ratio Rank (GRR), Mutual Information Rank
(MIR), Correlation Based Rank (CR), Conditional Entropy Rank (CER)~\cite{guyon_jmlr03}, Correlation-based Feature Subset Selection (CS)~\cite{Hall1998}, Reliff (RF)~\cite{Robnik-Sikonja2003} , Minimum Redundancy Maximum Relevancem(MRMR)~\cite{peng2005} and our proposed methods:  Disjunction Voting Expectation (DVE) and  Disjunction Voting likelihood (DVL), Conjunction Voting Expectation (CVE), Conjunction Voting likelihood (CVL). The first five methods are variable rank methods and the last seven are subset rank methods. The first five methods are variable rank methods that select variables by ranking them with some metric and the last seven are subset rank methods that assess subsets of variables and consider previous selections to decide the next selection.

@@

Next we examine Figure~\ref{fig:perf_vs_fs_alg}, where we evaluate the
performance distribution of various feature selection algorithms for
two datasets -- NG and S -- and classifier types -- Logistic
Regression and Naive Bayes -- at each stage of feature selection.
This graph shows that our CVL and CVE high recall feature selection
approaches provide strong recall performance at all stages of
feature selection in comparison to other selection methods.
%whose performance varies widely.
%We conjecture the reason for this stems from the strong agreement of
%features provided by the conjunctive voting scheme that helps reduce noise
%during feature selection process.

### 2014-02-05 trunk/ feature-selection-high-recall/FeatureSelectionProject/doc/fs_hr/intro.tex
Update

While a wide variety of feature selection methods have been proposed
in the literature, little research seems to focus on feature selection
specifically targeted to improve recall (minimization of false
negatives) in binary classification.  One reason for this is simply
that most feature selection methods are agnostic to the particular
supervised learning task -- applying to tasks from classification to
regression -- and hence are not focused on performance properties
specific to binary classifiers.  Yet, recall is an important aspect of
many binary classification problems (e.g., minimizing false negatives
in the identification of cancerous tumors) and it is critical to have
feature selection algorithms that are targeted to maintain high recall.

Naturally though, it would not make sense for a classifier or feature
selection method to focus on recall alone since the classifier which
always predicts \emph{true} (independent of the data) obtains optimal
recall performance.  To trade off precision with recall in order to
maintain high accuracy, one often uses a geometric average of the two
instead (known as F-score) but in practice most classifiers directly
optimize accuracy or some surrogate, e.g., a convex surrogate of 0-1
loss as in the SVM or maximum (conditional) likelihood as in Naive
Bayes or logistic regression.  So how can one achieve a high recall
SVM, Naive Bayes, or logistic regression classifier that already has a
well-defined accuracy-focused optimization criterion?  The idea we
pursue in this paper is that feature selection can help modulate the
recall performance of existing classifiers by encouraging selection of
features that cover more of the true cases in the data (hence
discouraging false negatives).

In this paper we propose a novel probabilistic voting model of feature
selection with the intent of encouraging false negative reduction
while still focusing on accuracy.  We argue that choosing features so
as to maximize likelihood objectives w.r.t. this model provides an
effective method for high recall feature selection.  Specifically, for
each objective, we derive an efficient formula for feature selection
in the filtering framework (i.e., greedy forward selection) and we
empirically compare the resulting feature selection criteria to a wide
variety of existing methods.

Our results demonstrate that our high recall feature selection
approach does indeed improve recall with improvements noticeable when
there is a high feature to data ratio.  Such results provide a novel
and efficient feature selection algorithm to target high-recall
classification tasks.
% TODO: elaborate more

>>>>

While a wide variety of feature selection methods have been proposed
in the literature, little research seems to focus on feature selection
specifically targeted to improve recall (minimization of false
negatives) in binary classification.  One reason for this is simply
that most feature selection methods are agnostic to the particular
supervised learning task -- applying to tasks from classification to
regression -- and hence are not focused on performance properties
specific to binary classifiers.  Yet, recall is an important aspect of
many binary classification problems (e.g., minimizing false negatives
in the identification of cancerous tumors) and it is critical to have
feature selection algorithms that can target features which provide
high recall coverage of the data.

Naturally though, it would not make sense for a classifier or feature
selection method to focus on recall alone since the classifier which
always predicts \emph{true} (independent of the data) obtains optimal
recall performance.  To trade off precision with recall in order to
maintain high accuracy, one often uses a geometric average of the two
instead (known as F-score) but in practice most classifiers directly
optimize accuracy or some surrogate, e.g., a convex surrogate of 0-1
loss as in the SVM or maximum (conditional) likelihood as in Naive
Bayes or logistic regression.  So how can one achieve a high recall
SVM, Naive Bayes, or logistic regression classifier that already has a
well-defined accuracy-focused optimization criterion?  The idea we
pursue in this paper is that feature selection can help modulate the
recall performance of existing classifiers by encouraging selection of
features that cover more of the true cases in the data (hence
discouraging false negatives harmful to recall).

In this paper we propose a novel probabilistic voting model of feature
selection with the intent of encouraging false negative reduction
while still focusing on accuracy.  We argue that choosing features so
as to maximize likelihood objectives w.r.t. this model provides an
effective method for high recall feature selection.  Specifically, for
each objective, we derive an efficient formula for feature selection
in the filtering framework (i.e., greedy forward selection) and we
empirically compare the resulting feature selection criteria to a wide
variety of existing filtering methods.

Our results demonstrate that our high recall feature selection
approach does indeed improve recall with improvements most significant when 
there is a high feature to data ratio.  Such results provide a novel
and efficient feature selection algorithm to target high-recall
classification tasks.
% TODO: elaborate more
